---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `gss`. Delete this note when before you submit 
your work. 

```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data

Sample: Data for this sample is gathered through personal interviews conducted on random members of the population. The data samples is from 1972 to 2012. 

This data set can be used for inference, but cannot be used to make causal conclusions or for causality studies.

The target population of the GSS is people that live in households in the U.S. over the age of 18. The survey is conducted in English and started being conducted in Spanish in 2006. This survey only targets those that live households.

Conclusions can only be made for the target population. 

* * *

## Part 2: Research question

<strong>Research Quesiton: </strong>

Is a connection between a person's region of residence and chances of living in poverty at the age of 16? 

As this survey covers 40 years of data and there have been many region changes in that time, a subset of time will need to be chosen to analyze. I decided to choose years greater than or equal to 2000, as this was a time relative economic prosperity for the U.S. and is more likely to point out the regions that are in trouble perpetually.

This question is of importance as it it shows which areas in the U.S. are in need of the most help, and where the system has failed. There are many negative effects to areas that have a high poverty level such as high crime and drugs. 

Important varables to consider in asnwering this question are: 
region — Region of interview.
year — GSS year for this respondent
incom16 — Rs family income when 16 years old

2. Poverty level by education level. 

3. Poverty level by race. 

```{r}
#set for consideration

r_i_study <- select(gss,year,region,incom16) %>% na.omit() %>%
  mutate(poverty=grepl("Far Below Average",incom16)) %>%
  mutate(y2000=as.factor(ifelse(year>=2000,"Y","N")))
```

* * *

## Part 3: Exploratory data analysis

Do start with I will output a table of all those that fit within the paramerters specified, since the start of the survey. 

```{r}
poverty_by_region_table <- table(r_i_study$region,r_i_study$poverty)
addmargins(poverty_by_region_table)
```

```{r}
prob_tab <- prop.table(poverty_by_region_table)
addmargins(prob_tab)
```

It can be seen in this table that the region with the highest poverty rate is the South Atlantic. The overall poverty rate of the sampled population is 8.5%, for the time period specified. 

It's easier to see visually in this bar graph.

```{r}
ggplot(r_i_study) + aes(x=region,fill=poverty) + geom_bar(position = "fill") +
  labs(x="Respondent's Region",y="Proportion",title="Povertly Level by Region") 
```

Next I will look at the data for the time period from 2000 to 2012. Which shows more modern data and distribution of poverty.  

```{r}
recent_povert <- filter(r_i_study,y2000=="Y")
poverty_by_region_table <- table(recent_povert$region,recent_povert$poverty)
addmargins(poverty_by_region_table)
addmargins(prop.table(poverty_by_region_table))
```

The poverty level at the age 16 of the sample population, increases from 2000-2016, it goes up to 9.9%. It would be worth investigating why the poverty level appears to be increasing. The South Atlantic is still the highest poverty level, in the U.S. only it has increased over this time span. 




* * *

## Part 4: Inference

<strong>Hypothesis</strong>

Null hypothesis: a respondents poverty level and region are independent variables. 
Alternative hypothesis: poverty level does indeed vary by region. 

The variables being used for this questions are categorical. The method for testing that my hypothesis given that categorical variable is being used is the chi-squared test of independence. 

In order to conduct a chi-squared test certain conditions must be met: 

Sampled data must be independent. This is true based on the sampling methods. 
There needs to be a large enough sample size, which is true, as the sample size is in the thousands. 


```{r}
chisq.test(r_i_study$region,r_i_study$poverty)
```


```{r}
chisq.test(recent_povert$region,recent_povert$poverty)
```

The test on the stat from 1972-2012 and the recent stat from 2000-2012 both produce p-values below alpha .05. This is reason to rejust the null hypothisis. 

This means that poverty levels do indead vary by region. 

Based on the results of this test further data could be gathered to determine why certain areas have greater poverty levels than others. 

